Quantile Regressions via Method of Moments with multiple fixed effects

### 2023-10-26 article/article.qmd
s
s

## Quantile Regression: Location-Scale model

@@

## Standard Errors: GLS, Robust, Clustered

@@

@@

>>>>

## Quantile Regression: Location-Scale model {#sec-betas}

@@

## Standard Errors: GLS, Robust, Clustered {#sec-se}

@@

Using the setup described in the previous section, @mss2019 proposes an extension to the model proposed by @he1997 that would allow for the estimation of quantile regression models with panel data, allowing for the inclusion of individual fixed effects. However, as the authors suggest, the methodology can be generalized to allow for the inclusion of multiple fixed effects. This type of analysis may be useful when considering data such as employer-employee linked data [@abowed2006], or teacher-student linked data [@harrissass2011]. Or, in the most common case, allowing to control for both individual and time fixed effects.

Reconsider the original model, and assume there are sets of unobserved heterogeneity that are assumed to be constant across observations, if they belong to common groups. In panel data, the groups would be the individual fixed effects and the time fixed effects. Without loss of generality, we can assume that the data generating process is as follows:

$$\begin{aligned}
  y_{i} &= x_{i}' \beta + \delta_{g1} + \delta_{g2} + \nu_i \\
  \nu_i &= \varepsilon_i \times (x_{i}' \gamma + \zeta_{g1} + \zeta_{g2})   
  \end{aligned}
$$

where we assume $x_{i}$ vary across groups $g_1$ and $g_2$, thus are not collinear, and that $\delta's$ and $\zeta's$ are the location and scale effects associated with groups fixed effects.[^4]

[^4]: We could just as well consider multiple sets of fixed effects

If the dimension of groups $g_k$ is low, this model could be estimated using a dummy inclussion approach following @sec-betas, and standard errors obtained as discussed in @sec-se. However, if the dimensionality of $g_k$ is high, the dummy inclusion approach may not be feasible. Instead, a more feasible approach is to apply the Frisch-Waugh-Lovell (FWL) theorem, and partial out the impact of the group fixed effects on the control variables $x_{i}$, the outcome of interest $y_{i}$, with a similar approach for the identification of $\sigma(x)$. In the case of unbalanced setups, with multiple groups, the estimation involves iterative processes for which various approaches have been suggested and implemented (see @correia_feasible_nodate, @gaure2013, @rios2015, among others).

When applying the partialing out approach, some modifications to the approach described in @sec-betas are needed.

1. For all dependent and independent variables in the model ($w=y,x$), we partial out the group fixed effects, and obtain the centered-residualized variables:

$$\begin{aligned}
w_{i} &= \delta_{g1}^w + \delta_{g2}^w + u_{i}^w \\
w_{i}^{rc} &= E(w_{i}) + \hat{u}_{i}^w
\end{aligned}
$$

2. We estimate the location model using the centered-residualized variables:

$$y_{i}^{rc} = x_{i}^{rc'} \beta + nu_{i}
$$

3. Since $|\hat \nu_i|$ is the dependent variable for the scale model, we apply the partialling out and recentering to this expression ($|\hat \nu_i|^{rc}$), and use that to estimate the model:

$$|\hat\nu_{i}|^{rc} = x_{i}^{rc'} \gamma + e_{i}$$

4. Finally the standardized residuals $\varepsilon_i$ can be obtained as follows

$$\hat{\varepsilon}_{i} = \frac{\nu_{i}}{|\hat\nu_{i}|- \hat e_{i}}$$

where $|\hat\nu_{i}|- \hat e_{i}$ is the prediction for the conditional standard deviation $\sigma(x_i)=x_{i}' \gamma + \zeta_{g1} + \zeta_{g2}$

The $\tau_{th}$ quantile of the error $\varepsilon$ can be estimated as usual, and the variance-covariance matrices obtained in the same way as before (see @sec-se), but using  $x_{i}^{rc}$ instead of $x_{i}$ when estimating the influence functions for all estimated coefficients.

@@

The estimation of quantile regression via moments assumes that the DGP is linear in parameters, with an heteroskedastic error term that is also linear function of parameters:

$$\begin{aligned}
y = x\beta + \nu \\
\nu = \varepsilon \times x\gamma
\end{aligned}
$$

where $\varepsilon$ is an unobserved i.i.d. random variable that is independent of $x$, and such that $x\gamma$ is larger than zero for any $x$.

In this case, the $\tau_{th}$ conditional quantile model can be written as:

$$q(y|\tau,X) = x(\beta + q(e|\tau) \times \gamma)$$

This model is identified under the following conditions:

$$\begin{aligned}
  E[(y_i-x_i'\beta)x_i ]  &= E[h_{1,i}]=0 \\
  E[ (|y_i-x_i'\beta|-x_i' \gamma) x_i ] &=E[h_{2,i}]=0 \\
  E\left[  \mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta\geq  y_i  \right) - \tau \right] 
   &=E[h_{3,i}]=0 
\end{aligned}
$$

In this model, to estimate the variance-covariance matrix the set of coefficients $\theta'=[\beta' \ \gamma' \ q(\varepsilon|\tau)']$, we need to obtain the influence functions of all coefficients, which are defined as:

$$\lambda_i = \bar G(\theta)^{-1}
\begin{bmatrix}
h_{1,i} \\
h_{2,i} \\
h_{3,i}
\end{bmatrix}
$$

where the Jacobian matrix $\bar G(\theta)$ is defined as:

$$\bar G(\theta) = \begin{bmatrix}
\bar G_{11} & G_{12} & G_{13} \\
\bar G_{21} & \bar G_{22} & G_{23} \\
\bar G_{31} & \bar G_{32} & \bar G_{13} \\
\end{bmatrix}
$$

with

$$\bar G_{j,k} = - \frac 1 N \sum_{i=1}^N \frac{\partial h_{j,i}}{\partial \theta_k'} \ \forall j,k \in 1,2,3
$$

### First Moment Condition: Location Model

$$h_{1,i}=x_i(y_i-x_i'\beta)$$

$$\begin{aligned}
\bar G_{1,1} &=- \frac{1}{N} \sum_{i=1}^N \frac{\partial h_{1,i}}{\partial \beta'} \\
             &=- \frac{1}{N} \sum_{i=1}^N (-x_i x_i') \\
             &= N^{-1} X'X
\end{aligned}
$$

$$
\bar G_{1,2} = \bar G_{1,3} = 0
$$

### Second Moment Condition: Scale model

$$h_{2,i}=x_i(|y_i-x_i'\beta|-x_i'\gamma)$$

$$\begin{aligned}
\bar G_{2,1} &= -\frac{1}{N} \sum \frac{\partial h_{2,i}}{\partial \beta'} \\
             &=  \frac{1}{N} \sum x_i x_i' \frac{y_i-x_i'\beta}{|y_i-x_i'\beta|} \\
\frac{y_i-x_i'\beta}{|y_i-x_i'\beta|} &= sign(y_i-x_i'\beta) \\            
\end{aligned}
$$

Under the assumption $\varepsilon_i \times x\gamma$, or in this case $y_i-x_i'\beta$, is uncorrelated with $x$, we can simplify the expression as:

$$\begin{aligned}
\bar G_{2,1} &= N^{-1} \left(N^{-1}\sum sign(y_i-x_i'\beta)\right) \sum x_i x_i' \\
&= N^{-1} E[sign(y_i-x_i'\beta)] X'X
\end{aligned}
$$

$$\begin{aligned}
\bar G_{2,2} &= -\frac{1}{N} \sum \frac{\partial h_{2,i}}{\partial \gamma'} \\
             &=  \frac{1}{N} \sum x_i x_i' 
             &= N^{-1} X'X
\end{aligned}
$$

$$\bar G_{2,3}=0$$

### Third Moment Condition: Quantile of Standardized Residual

$$h_{3,i} = \mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i \geq 0 \right) - \tau$$

Because the indicator function $\mathbb{1}()$ is not differentiable, we impose some smoothing assumptions:

First:

$$\begin{aligned}
E[h_{3,i}] -\tau &= E[\mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i \geq 0 \right)] - \tau \\
                &= E\left[E[\mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta - y_i \geq 0 \right)|x]\right] - \tau
\end{aligned}
$$

Because we are aiming to identify the conditional quantile function of $y$, and under the assumption of the DGP it follows that:

$$\begin{aligned}
q(y|\tau) &= q(\varepsilon|\tau) x_i'\gamma +x_i'\beta \\
E[\mathbb{1}\left( q(y|\tau) - y_i \geq 0 \right)|x] &= \tau 
\end{aligned}
$$

and asymptotically, we expect that

$$E[\mathbb{1}\left( q(y|\tau) - y_i \geq 0 \right)|x]=
E[\mathbb{1}\left( q(y|\tau) \geq  y_i \right)|x] = \tau = F_{y |X} \big( q(y|\tau) \big) 
$$

but also that

$$F_{y|X} \big( q(y|\tau)) = \tau = F_{\varepsilon}( q(\varepsilon|\tau) )
$$

where $F_{y|X}$ is the conditional cumulative distribution function (CDF) of $y$, and $F_{\varepsilon}$ is the unconditonal CDF of the iid error $\varepsilon$.

Second, based on non-parametric estimation of the CDF, we can approximate $F_{y|X} \big( q(y|\tau))$ as:

$$F_{y|X} \big( q(y|\tau) ) \approx \frac{\sum_{i=1}^N F_{y|X) ( q(y|\tau) - y_i)

### 2023-10-26 article/article.tex
s
s

\title{Quantile Regressions via Method of Moments with multiple fixed
effects}
\author{Fernando Rios-Avila \and Leonardo Siles \and Gustavo
Canavire-Bacarreza}
\date{2023-10-24}

\begin{document}
\maketitle
\begin{abstract}
This paper proposes a new method to estimate quantile regressions with
multiple fixed effects. The method expands on the strategy proposed by
Machado and Santos Silva (2019), allowing for multiple fixed effects,
and providing various alternatives for the estimation of Standard
errors. We provide Monte Carlo simulations to show the finite sample
properties of the proposed method in the presence of two sets of fixed
effects. Finally, we apply the proposed method to estimate
\textbf{something interesting}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

@@

\hypertarget{quantile-regression-location-scale-model}{%
\subsection{Quantile Regression: Location-Scale
model}\label{quantile-regression-location-scale-model}}

@@

\hypertarget{standard-errors-gls-robust-clustered}{%
\subsection{Standard Errors: GLS, Robust,
Clustered}\label{standard-errors-gls-robust-clustered}}

@@

@@

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-cameron2005}{}}%
Cameron, Adrian Colin, and Pravin K. Trivedi. 2005.
\emph{Microeconometrics: Methods and Applications}. Cambridge ; New
York: Cambridge University Press.

@@

@@

@@

>>>>

\title{Quantile Regressions via Method of Moments with multiple fixed
effects}
\author{Fernando Rios-Avila \and Leonardo Siles \and Gustavo
Canavire-Bacarreza}
\date{2023-10-26}

\begin{document}
\maketitle
\begin{abstract}
This paper proposes a new method to estimate quantile regressions with
multiple fixed effects. The method expands on the strategy proposed by
Machado and Santos Silva (2019), allowing for multiple fixed effects,
and providing various alternatives for the estimation of Standard
errors. We provide Monte Carlo simulations to show the finite sample
properties of the proposed method in the presence of two sets of fixed
effects. Finally, we apply the proposed method to estimate
\textbf{something interesting}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, breakable]}{\end{tcolorbox}}\fi

@@

\hypertarget{sec-betas}{%
\subsection{Quantile Regression: Location-Scale model}\label{sec-betas}}

@@

\hypertarget{sec-se}{%
\subsection{Standard Errors: GLS, Robust, Clustered}\label{sec-se}}

@@

Using the setup describe in the previous section, Machado and Santos
Silva (2019) proposes an extension to the model proposed by He (1997)
that would allow for the estimation of quantile regression models with
panel data, allowing for the inclussion of individual fixed effects.
However, as the authors suggest, the methodology can be generalized to
allow for the inclusion of multiple fixed effects. This type of analysis
may be useful when considering data such as employer-employee linked
data (Abowd, Kramarz, and Roux 2006), or teacher-student linked data
(Harris and Sass 2011). Or, in the most common case, allowing to control
for both individual and time fixed effects.

Reconsider the original model, and assume there are sets of unobserved
heterogeneity that are assumed to be constant across observations, if
they belong to common groups. In panel data, the groups would be the
individual fixed effects and the time fixed effects. Without loss of
generality, we can assume that the data generating process is as
follows:

\[\begin{aligned}
  y_{i} &= x_{i}' \beta + \delta_{g1} + \delta_{g2} + \nu_i \\
  \nu_i &= \varepsilon_i \times (x_{i}' \gamma + \zeta_{g1} + \zeta_{g2})   
  \end{aligned}
\]

where we assume \(x_{i}\) vary across groups \(g_1\) and \(g_2\), thus
are not collinear, and that \(\delta's\) and \(\zeta's\) are the
location and scale effects associated with groups fixed
effects.\footnote{We could just as well consider multiple sets of fixed
  effects}

If the dimension of groups \(g_k\) is low, this model could be estimated
using a dummy inclussion approach following Section~\ref{sec-betas}, and
standard errors obtained as discussed in Section~\ref{sec-se}. However,
if the dimensionsionality of \(g_k\) is high, dummy inclussion approach
may not be feasible. Instead, a more feasible approach is to apply the
Frisch-Waugh-Lovell (FWL) theorem, and partial out the impact of the
group fixed effects on the control variables \(x_{i}\), the outcome of
interest \(y_{i}\), with a similar approach for the identification of
\(\sigma(x)\). In the case of unbalanced setups, with multiple groups,
the estimation involves iterative processes for which various approaches
have been sugested and implemented (see Correia (2016), Gaure (2013),
Rios-Avila (2015), among others).

When applying the partialing out approach, some modifications to the
approach described in Section~\ref{sec-betas} is needed.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  For all dependent and independent variables in the model (\(w=y,x\)),
  we partial out the group fixed effects, and obtain the
  centered-residualized variables:
\end{enumerate}

\[\begin{aligned}
w_{i} &= \delta_{g1}^w + \delta_{g2}^w + u_{i}^w \\
w_{i}^{rc} &= E(w_{i}) + \hat{u}_{i}^w
\end{aligned}
\]

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  We estimate the location model using the centered-residualized
  variables:
\end{enumerate}

\[y_{i}^{rc} = x_{i}^{rc'} \beta + nu_{i}
\]

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Since \(|\hat \nu_i|\) is the dependent variable for the scale model,
  we apply the partialling out and recentering to this expression
  (\(|\hat \nu_i|^{rc}\)), and use that to estimate the model:
\end{enumerate}

\[|\hat\nu_{i}|^{rc} = x_{i}^{rc'} \gamma + e_{i}\]

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Finally the standardized residuals \(\varepsilon_i\) can be obtained
  as follows
\end{enumerate}

\[\hat{\varepsilon}_{i} = \frac{\nu_{i}}{|\hat\nu_{i}|- \hat e_{i}}\]

where \(|\hat\nu_{i}|- \hat e_{i}\) is the prediction for the
conditional standard deviation
\(\sigma(x_i)=x_{i}' \gamma + \zeta_{g1} + \zeta_{g2}\)

The \(\tau_{th}\) quantile of the error \(\varepsilon\) can be estimated
as usual, and the variance-covariance matrices obtained in the same way
as before (see Section~\ref{sec-se}), but using \(x_{i}^{rc}\) instead
of \(x_{i}\) when estimating the influence functions for all estimated
coefficients.

@@

The estimation of quantile regression via moments assumes that the DGP
is linear in parameters, with an heteroskedastic error term that is also
linear function of parameters:

\[y = x\beta + e_i \times x\gamma\]

where \(e_i\) is an unobserved i.i.d. random variable that is
independent of \(x\), and such that \(x\gamma\) is larger than zero for
any \(x\).

In this case, the \(\tau_{th}\)conditional quantile model can be written
as:

\[q(y|\tau,X) = x(\beta + q(e|\tau) \times \gamma)\]

This model is identified under the following conditions:

\[\begin{aligned}
  E[(y_i-x_i'\beta)x_i ]  &= E[h_{1,i}]=0 \\
  E[ (|y_i-x_i'\beta|-x_i' \gamma) x_i ] &=E[h_{2,i}]=0 \\
  E\left[  \mathbb{1}\left( q(\varepsilon|\tau) x_i'\gamma +x_i'\beta\geq  y_i  \right) - \tau \right] 
   &=E[h_{3,i}]=0 
\end{aligned}
\]

In this model, to estimate the variance-covariance matrix the set of
coefficients \(\theta'=[\beta' \ \gamma' \ q(\varepsilon|\tau)']\), we
need to obtain the influence functions of all coefficients, which are
defined as:

\[\lambda_i = \bar G(\theta)^{-1}
\begin{bmatrix}
h_{1,i} \\
h_{2,i} \\
h_{3,i}
\end{bmatrix}
\]

where the Jacobian matrix \(\bar G(\theta)\) is defined as:

\[\bar G(\theta) = \begin{bmatrix}
\bar G_{11} & G_{12} & G_{13} \\
\bar G_{21} & \bar G_{22} & G_{23} \\
\bar G_{31} & \bar G_{32} & \bar G_{13} \\
\end{bmatrix}
\]

with

\[\bar G_{j,k} = - \frac 1 N \sum_{i=1}^N \frac{\partial h_{j,i}}{\partial \theta_k'} \ \forall j,k \in 1,2,3
\]

\hypertarget{first-moment-condition-location-model}{%
\subsubsection{First Moment Condition: Location
Model}\label{first-moment-condition-location-model}}

\[h_{1,i}=x_i(y_i-x_i'\beta)\]

\[\begin{aligned}
\bar G_{1,1} &=- \frac{1}{N} \sum_{i=1}^N \frac{\partial h_{j,i}}{\partial \beta'} \\
             &=- \frac{1}{N} \sum_{i=1}^N (-x_i x_i') \\
             &= N^{-1} X'X
\end{aligned}
\]

\[\begin{aligned}
\bar G_{1,2} = \bar G_{1,3} = 0
\]

\hypertarget{second-moment-condition-scale-model}{%
\subsubsection{Second Moment Condition: Scale
model}\label{second-moment-condition-scale-model}}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abowed2006}{}}%
Abowd, John M., Francis Kramarz, and Sébastien Roux. 2006. {``Wages,
Mobility and Firm Performance: Advantages and Insights from Using
Matched Worker--Firm Data*.''} \emph{The Economic Journal} 116 (512):
F245--85.
https://doi.org/\url{https://doi.org/10.1111/j.1468-0297.2006.01099.x}.

\leavevmode\vadjust pre{\hypertarget{ref-cameron2005}{}}%
Cameron, Adrian Colin, and Pravin K. Trivedi. 2005.
\emph{Microeconometrics: Methods and Applications}. Cambridge ; New
York: Cambridge University Press.

@@

\leavevmode\vadjust pre{\hypertarget{ref-correia_feasible_nodate}{}}%
Correia, Sergio. 2016. {``A {Feasible} {Estimator} for {Linear} {Models}
with {Multi}-{Way} {Fixed} {Effects}.''}
\url{http://scorreia.com/research/hdfe.pdf}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-gaure2013}{}}%
Gaure, Simen. 2013. {``{OLS} with Multiple High Dimensional Category
Variables.''} \emph{Computational Statistics \& Data Analysis} 66:
8--18. https://doi.org/\url{https://doi.org/10.1016/j.csda.2013.03.024}.

\leavevmode\vadjust pre{\hypertarget{ref-harrissass2011}{}}%
Harris, Douglas N., and Tim R. Sass. 2011. {``Teacher Training, Teacher
Quality and Student Achievement.''} \emph{Journal of Public Economics}
95 (7): 798--812.
https://doi.org/\url{https://doi.org/10.1016/j.jpubeco.2010.11.009}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-rios2015}{}}%
Rios-Avila, Fernando. 2015. {``Feasible {Fitting} of {Linear} {Models}
with {N} {Fixed} {Effects}.''} \emph{The Stata Journal} 15 (3): 881--98.
\url{https://doi.org/10.1177/1536867X1501500318}.
